---
title: Event Processing
author: andrea perlato
date: '2019-01-23'
slug: event-processing
categories:
  - theory
tags:
  - event processing
---

<style>
body {
text-align: justify}
</style>

**Process Data** </br>
Event data consists of three basic components: the why, the what and the who. </br>
Analysing event data is an iteractive process of three steps: extraction (from raw data to event log), processing (removing redundant details, enrich data by calculating variables) and analysis. </br>
The analysis could be for instance which are the roles of different doctors and nurses **organization** and how they work together. There is also in the analysis the **controll-flow prospective** (e.g. what is the journey of a patient throughthe emergency rooms). Finally, the performance prospective, focus on the time and **efficiency** (e.g. how long does it take before a patient can leave the emergency rooms). </br>

**A first glimpse of the event log**
Some initial question about event log are: </br>
How many **cases** are described? </br>
How many distinct **activities** are performed? </br>
How many **events** are recorded? </br>
What is the **time period** in which the data is recorded? </br>

To analys the process data, we use the [**bupaR package**](https://www.bupar.net/index.html).
In the following exercise we will have a look at a process data, which descries the journey of patients in a hospital.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
library(bupaR)
# n_cases(patients) # how many patients
# summary(patients) 
slice(patients, 1) # show the first patient

```

As shown above, the first patient went through six events: registration, triage, blood tests, an MRI scan, discussion of results, and check out, over the course of a week in January 2017. </br>

Here below, we see that the most common activities are administrative: registration (18.4%), triage (18.4%), discussing results (18.2%), and checking out (18.1%). </br>

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
activities(patients)

```

The sequence of activities performed in relation to one case is called its **trace**, literally the trace that the process instance leaves in our data.
The trace plot here below, shows two **happy paths** ( most frequebt sequences: 51.6% and 46.8% of cases). They start with registration and triage, then have one or more treatments, then end with discussion of results and check-out.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '80%'}
library(processmapR)
trace_explorer(patients, coverage = 1)

```

Another way to visualize processes is by constructing a **process map**. </br>
A process map is a **directed graph** that shows the **activities** of the process and the flows between them. The colors of the nodes and the thickness of the arrows indicate the most frequent activities and process flows.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
process_map(patients) # draw the process map

```
The process map is a great way to see which activities each happy path has in common.
In this article, we discovered a methodology for analyzing process data, consisting of three stages: extraction, processing and analysis. There are the specific elements of process data which are required for analysis.    

**Resource activity matrix** </br>
The nest step is to look at the activities. In the resource activity matrix we put different resources on one dimention and the activities on the other dimension. </br>
In the following example, we will look at the employees working in a hospital which are treating patients with [**Sepsis**](https://en.wikipedia.org/wiki/Sepsis), a life-threatening condition caused by an infection. Treating these patients is of the utmost importance. Let's see how tasks are divided. 

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
# Calculate frequency of resources
frequencies_of_resources <- resource_frequency(sepsis, level = "resource-activity")

# See the result as a plot
plot(frequencies_of_resources)

```
In the plot above, we can see which resources executed which activities, and how often.
Resources F,G,I,M and O have the same task; they are responsible for the activity Admission NC. </br>
Momreover, the most common combination was Resource B performing the Leucocytes activity. Let's see what this information tells us about the employees of the hospital. </br>
Another thing that can be measured is their **involvement**. </br>
Involvement refers to the number of cases a resource works on, irrespective of his/her total workload.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
# Calculate resource involvement
involvement_of_resources <- resource_involvement(sepsis, level = "resource")

# Show the result in a plot
plot(involvement_of_resources)

```

**Control Flow** </br>
It refers to the different successiones of activities. Each case can be expressed as a sequence of activities. Each unique sequence is called a trace (or process variant). There are several ways to look at a trace in a process. On the one hand, we have metrics (enty and exit poin, lengh of cases, presence of activities, rework). On the other hand, we have various tools to look at control-flow patterns (process map, trace explorer, precedence matrix).

**Rework** </br>
When looking at the structure of a process, an interesting aspect to look at is rework. Rework happens when the same activity is repeated within a single case, which often points towards time and resources which are wasted, and typically a mistake which has to be rectified. </br>
We can distinguish between two different types of rework: **immediate rework**, we are also known as self-loops of activities, and **rework which surfaces later in the process**, which we call repetitions.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
# Number of repetitions per activity
n_reps_per_activity <- number_of_repetitions(sepsis, level = "activity") 

# See the result in a plot
plot(n_reps_per_activity)

```

The CRP activity was repeated more than any other activity type. </br>
















