---
title: Event Processing
author: andrea perlato
date: '2019-01-23'
slug: event-processing
categories:
  - theory
tags:
  - event processing
---

<style>
body {
text-align: justify}
</style>

**Process Data** </br>
Event data consists of three basic components: the why, the what and the who. </br>
Analysing event data is an iteractive process of three steps: extraction (from raw data to event log), processing (removing redundant details, enrich data by calculating variables) and analysis. </br>
The analysis could be for instance which are the roles of different doctors and nurses **organization** and how they work together. There is also in the analysis the **controll-flow prospective** (e.g. what is the journey of a patient throughthe emergency rooms). Finally, the performance prospective, focus on the time and **efficiency** (e.g. how long does it take before a patient can leave the emergency rooms). </br>

**A first glimpse of the event log**
Some initial question about event log are: </br>
How many **cases** are described? </br>
How many distinct **activities** are performed? </br>
How many **events** are recorded? </br>
What is the **time period** in which the data is recorded? </br>

To analys the process data, we use the [**bupaR package**](https://www.bupar.net/index.html).
In the following exercise we will have a look at a process data, which descries the journey of patients in a hospital.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
library(bupaR)
# n_cases(patients) # how many patients
# summary(patients) 
slice(patients, 1) # show the first patient

```

As shown above, the first patient went through six events: registration, triage, blood tests, an MRI scan, discussion of results, and check out, over the course of a week in January 2017. </br>

Here below, we see that the most common activities are administrative: registration (18.4%), triage (18.4%), discussing results (18.2%), and checking out (18.1%). </br>

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
activities(patients)

```

The sequence of activities performed in relation to one case is called its **trace**, literally the trace that the process instance leaves in our data.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '80%'}
library(processmapR)
trace_explorer(patients, coverage = 1)

```









