---
title: Distinguish Benign and Malign Tumor via ANN
author: andrea perlato
date: '2019-02-05'
slug: distinguish-benign-and-malign-tumor-via-ann
categories:
  - artificial intelligence
tags:
  - ANN
---

<style>
body {
text-align: justify}
</style>

We try to recognize cancer in human breast using a multi-hidden layer artificial neural network via H2O package.
We use the Wisconsin Breast-Cancer Dataset which is a collectioin of Dr.Wolberg real clinical cases. There are no images, but we can recognize malignal tumor based on 10 biomedical attributes. We have a total number of 699 patients divided in two classes: malignal and benign cancer.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
library(mlbench)
library(h2o)
h2o.init(nthreads = -1) # initializing h2o

```

The table below shows the crucial biomedical features involved in Cancer.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
library(knitr)
library(kableExtra)
library(formattable)

data("BreastCancer")
dt <- as.data.frame(BreastCancer)
dt <- dt[1:10,]

kable(dt) %>%
  kable_styling(bootstrap_options = "responsive", full_width = T, position = "center", font_size = 16)


```








