---
title: How to create 3D and 4D plot
author: andrea perlato
date: '2019-06-07'
slug: how-to-create-3d-and-4d-plot
categories:
  - graph
tags:
  - 3D
  - 4D
---

<style>
body {
text-align: justify}
</style>

A 3D plot is quite popular, in particular in business presentation, but it is almost always inappropriately used. In fact, it is rare to see a 3D plot that could not be improved by turning into a regular 2D figure. </br>
Visualizations using 3D position scales can sometimes be appropriate, however. If the visualization is show it slowly rotation, rather than a static image from one prospective, will allow the viewer to discern where in 3D space different graphicla elements resides. The human brain is very good at reconstructing a 3D scene from a series of images taken from different angles, and the slow rotation of the graphic provides exactly these images. </br>
We’ll use the [**iris data set**](https://en.wikipedia.org/wiki/Iris_flower_data_set) in the following examples.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
# Load data
data(iris)
head(iris)

# Load libraries
library(plot3D)
library(rgl)
library(plot3Drgl)

# Define coordinates
x <- sep.l <- iris$Sepal.Length
y <- pet.l <- iris$Petal.Length
z <- sep.w <- iris$Sepal.Width

# Add small dots on basal plane and on the depth plane
scatter3D_fancy <- function(x, y, z,..., colvar = z)
  {
   panelfirst <- function(pmat) {
      XY <- trans3D(x, y, z = rep(min(z), length(z)), pmat = pmat)
      scatter2D(XY$x, XY$y, colvar = colvar, pch = ".", 
              cex = 2, add = TRUE, colkey = FALSE)
   
      XY <- trans3D(x = rep(min(x), length(x)), y, z, pmat = pmat)
      scatter2D(XY$x, XY$y, colvar = colvar, pch = ".", 
              cex = 2, add = TRUE, colkey = FALSE)
  }
  scatter3D(x, y, z, ..., colvar = colvar, panel.first=panelfirst,
    colkey = list(length = 0.5, width = 0.5, cex.clab = 0.75)) 
}

# Define like color
scatter3D_fancy(x, y, z, pch = 16,
    ticktype = "detailed", theta = 15, d = 2,
    main = "Iris data",  clab = c("Petal", "Width (cm)"))


```

In order to generate a fancy scatter plot with small dots on balsal palne the scater3D_fancu was created. We will create a simila function for histograms.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
hist3D_fancy<- function(x, y, break.func = c("Sturges", "scott", "FD"), breaks = NULL,
                        colvar = NULL, col="white", clab=NULL, phi = 5, theta = 25, ...){
  
# Compute the number of classes for a histogram
  break.func <- break.func [1]
  if(is.null(breaks)){
    x.breaks <- switch(break.func,
                       Sturges = nclass.Sturges(x),
                       scott = nclass.scott(x),
                       FD = nclass.FD(x))
    y.breaks <- switch(break.func,
                       Sturges = nclass.Sturges(y),
                       scott = nclass.scott(y),
                       FD = nclass.FD(y))
  } else x.breaks <- y.breaks <- breaks
  
# Cut x and y variables in bins for counting
  x.bin <- seq(min(x), max(x), length.out = x.breaks)
  y.bin <- seq(min(y), max(y), length.out = y.breaks)
  xy <- table(cut(x, x.bin), cut(y, y.bin))
  z <- xy
  
  xmid <- 0.5*(x.bin[-1] + x.bin[-length(x.bin)])
  ymid <- 0.5*(y.bin[-1] + y.bin[-length(y.bin)])
  
  oldmar <- par("mar")
  par (mar = par("mar") + c(0, 0, 0, 2))
  hist3D(x = xmid, y = ymid, z = xy, ...,
    zlim = c(-max(z)/2, max(z)), zlab = "counts", bty= "g", 
    phi = phi, theta = theta,
    shade = 0.2, col = col, border = "black",
    d = 1, ticktype = "detailed")
   
  scatter3D(x, y,
    z = rep(-max(z)/2, length.out = length(x)),
    colvar = colvar, col = gg.col(100),
    add = TRUE, pch = 18, clab = clab,
    colkey = list(length = 0.5, width = 0.5,
       dist = 0.05, cex.axis = 0.8, cex.clab = 0.8)
       )
  par(mar = oldmar)
}


# Create his3D using plot3D
# hist3D_fancy(iris$Sepal.Length, iris$Petal.Width, colvar=as.numeric(iris$Species))

# Make the rgl version
# plotrgl()

```



<center>
![](/img/histogram3d){width=100%}
</center>















