---
title: Interactive Tables
author: andrea perlato
date: '2019-03-14'
slug: interactive-tables
categories:
  - graph
tags:
  - interactive tables
---

<style>
body {
text-align: justify}
</style>

Oftern, it is useful to provide interactive tables alonside charts.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
library(tidyverse)
library(DT)
transport_data <- read_csv("C:/07 - R Website/dataset/Graph/transport_data.csv")

# Column filtering
datatable(transport_data,
          rownames = FALSE,
          extensions = c("Responsive"))

```

As we can see from the example above, we can horizonatally and vertically scoll through the data. We can also filter globally the data.
At this point, We can improve the readability of the table, and add a filter at the top of each columns. We can also manipulate the pagination of the table. For example, in the example below we set the option page lenght to 5 instead of 10.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
pretty_headers <- gsub("[.]"," ", colnames(transport_data)) %>%
  str_to_title()

transport_data %>%
  datatable(rownames = FALSE,
            colnames = pretty_headers,
            filter = list(position = "top"),
            options = list(pageLength = 5),
            extensions = c("Responsive"))

```







