---
title: Interactive Forecasting
author: andrea perlato
date: '2019-04-03'
slug: interactive-forecasting
categories:
  - graph
  - time series
tags:
  - forecasting
---

<style>
body {
text-align: justify}
</style>

In this post, we use some fairly new technology of time series analysis namely neural nets and interactive charting tools.

```{r, echo=FALSE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '80%'}
library(knitr)
library(kableExtra)
library(formattable)
revenue <- read.csv("C:/07 - R Website/dataset/TS/camping-revenue-97-17.csv", sep = "," , quote = "")
colnames(revenue)[1] <- "V1"
colnames(revenue)[2] <- "V2"

# Clean data
library(tidyr)
revenue <- separate(revenue, col = V2,
                    sep = c(2, -3),
                    into = c("resr", "data", "rest2"))


# Convert to time series
myts <- ts(as.numeric(revenue$data), 
           start = 1997, frequency = 12)

library(forecast)
myts <- tsclean(myts)

```

**INTERACTIVE GRAPH** </br>
The time series results should be presented interactively in order to highlight certain features. 

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
# Handle outliers
library(forecast)
myts <- tsclean(myts)

# Set up a NN
mynnetar <- nnetar(myts)

# Forecast 3 years
nnetforecast <- forecast(mynnetar, h = 36, PI = TRUE) # PI create the prediction intervals for the forecast

library(ggplot2)
# Data we need for the graph
data <- nnetforecast$x # raw data
lower <- nnetforecast$lower[,2] # confidence intervals lower bound
upper <- nnetforecast$upper[,2] # confidence intervals upper bound
pforecast <- nnetforecast$mean # th element mean

mydata <- cbind(data, lower, upper, pforecast) # put everything in one dataframe

library(dygraphs)
dygraph(mydata, main = "Campsite Restaurant") %>%                                       # get data and the caption
  dyRangeSelector() %>%                                                                 # the zoom tool
  dySeries(name = "data", label = "Revenue Data") %>%                                   # add time series which are store in: data <- nnetforecast$x
  dySeries(c("lower","pforecast","upper"), label = "Revenue Forecast") %>%              # add the forecast and CI
  dyLegend(show = "always", hideOnMouseOut = FALSE) %>%                                 # add the legend (time series + forecast)
  dyAxis("y", label = "Monthly Revenue USD") %>%                                        # label the y axis
  dyHighlight(highlightCircleSize = 5,                                                  # specify what happen when the mouse in hovering the graph
              highlightSeriesOpts = list(strokeWidth = 2)) %>%
  dyOptions(axisLineColor = "navy", gridLineColor = "grey") %>%                         # set axis and fridline colors
  dyAnnotation("2010-8-1", text = "CF", tooltip = "Camp Festival", attachAtBottom = T)  # add annotation

```

The Interactive graph above consists in two distinct fractions. The longer line at the beginning is the original data. The second part which is 36 months long gets activated as soon as we hover it, and this is the forecasting with prediction intervals. </br>
At the bottom we have an interactive range selector where we can specify the level of zoom in and out in the timeline. </br>
To lear more go to [**Neural Network Autoregressive Model NNAR**](https://andrea-perlato.netlify.com/tspost/neural-nets-and-interactive-graphs/)





















