---
title: Quantile Regression in Medical Expenditures
author: andrea perlato
date: '2019-03-20'
slug: quantile-regression-in-medical-expenditures
categories:
  - machine learning
tags:
  - quantile regression
---

<style>
body {
text-align: justify}
</style>

The Quantile regression gives a more comprehensive picture of the effect of the independent variables on the dependent variable.  Instead of estimating the model with average effects using the OLS linear model, the quantile regression produces different effects along the distribution (quantiles) of the dependent variable.  The dependent variable is continuous with no zeros or too many repeated values. </br>
Examples include estimating the effects of household income on food expenditures for low- and high-expenditure households, what are the factors influencing total medical expenditures for people with low, medium and high expenditures. </br>
The following oexample is based on **Medical Expenditure Panel Survey MEPS**. The dependent variable is the **total medical expenditures**, and the independent variables are: **supplemental insurance**, **total number of chronic conditions**, **age**, **female**, and **white**. </br>
We estimate an OLS regression, and quantile regression at 25th, 50th, and 75th quantile. </br>
The standard **Ordinary Least Squares OLS**  models the relationship between one or more independent variables and the conditional mean of a dependent variable. The **Quantile Regression** models the relationship betwwn the conditional quantiles rather than just the conditional mean of the dependent variable. A quantile regression gives a more comprehensive picture of the effect of the independent variables on the dependent variable because we can show different effects (quantiles). </br>
One pratical consideration is that the distribution of the dependent variable has to be continuous and it shouldn't has zero or too many repeated values. </br>
One important aspect to take in considertion in Quantile Regression is that coefficients can be significanlty different than the OLS coefficients, showing different effects along the distribution of the dependent variable. </br>
The advantages of the Quantile regression are: </br>
Flexibility for modeling data with heterogeneous conditional distributions. </br>
Median regression is more robust to outliers than the OLS regression. </br>
Quantile regression can show different effects of the independent variables on the dependent variable depending across the spectrum of the dependent variable. </br>

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '80%'}
library(quantreg)
mydata <- read.csv("C:/07 - R Website/dataset/ML/quantile_health.csv")
attach(mydata)
summary(mydata)

# Define variables
Y <- cbind(totexp)
X <- cbind(suppins, totchr, age, female, white)

```

The variable **totexp** is the total expenditure and is dependent variable. The independent variables are **suppins** supplemental insurance, **totchr** total number of chronic conditions, **age**, **female**, and **white**. </br>
The first step is to perform an OLS regression.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '80%'}
# OLS regression
olsreg <- lm(Y ~ X, data=mydata)
summary(olsreg)

```

The **Xtotchr**, the total of chronic condition, says that each of chronich condition brings 2528.079 more dollars in **totexp** total medical expenditure. </br>
Now, we perform a quantile regression.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '80%'}
# Simultaneous quantile regression
quantreg2575 <- rq(Y ~ X, data=mydata, tau=c(0.25, 0.75))
summary(quantreg2575)

```
The **Xtotchr**, the total of chronic condition, for the **0.25 quantile** is 782.47, and the interpretatation is: adding 25th quantile each of chronich condition brings only 782.42 more dollars in **totexp** total medical expenditure. This is a much ower value that we had before with OLS. 
This means, for low number of chronic conditions the medical expenditure is lower. </br>
On the oder hand, looking at the **0.75 quantile** for the total of chronic condition we have 2855.31 more dollar per each more chronic condition. This value is more similar with the OLS coefficient, and in fact this time we have not a significant difference from the OLS coefficient. </br>

We can also perform an ANOVA to compare the coefficient at 25th quantile vs. 75th quantile.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '80%'}
# Quantile regression at 25 the quanile
quantreg25 <- rq(Y ~ X, data=mydata, tau=0.25)

# Quantile regression at 75 the quanile
quantreg75 <- rq(Y ~ X, data=mydata, tau=0.75)

# ANOVA test for coefficient differences
anova(quantreg25, quantreg75)

```

As we can see from the result above, there is a significant difference in the coefficients, and this justify to use the quantile regression. </br>
Now, we can plot the data and the coefficinets we found from the quantile regression.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '80%'}
# Plotting data
quantreg.all <- rq(Y ~ X, tau = seq(0.05, 0.95, by = 0.05), data=mydata)
quantreg.plot <- summary(quantreg.all)
plot(quantreg.plot)

```

We focus our attention on the **Xtotchr** (total of chronic condition) graph. </br>
The red orizontal line is the OLS coefficient, and we can see that the value is exactly the same of what we found before (2528.079). Notice that the OLS line is flat along the quantile in the x-axis, because it cannot vary across the quantiles.
Looking at the quantile trend (black curve with grey confidence intervals), we can see that for low quantiles there is a significant difference below OLS. On the contrary, there is a significant difference above OLS for high quantile. </br>
Again, we can see from the graph of **Xtotchr** that there is not a significant difference for the 75th quantile. </br>

Looking at the **Xage** graph, there is not a significat difference from OLS across the quantiles, except at the last quantile.




















