---
title: Generalized Addictive Models GAMs
author: andrea perlato
date: '2019-03-22'
slug: generalized-addictive-models-gams
categories:
  - machine learning
tags:
  - gams
  - generalized addictive models
---

<style>
body {
text-align: justify}
</style>

Generalized Addictive Models GAMs incorporates non linear form of predictions, and are useful when we have not linearity between response variable and predictors. [**GAMs**](https://en.wikipedia.org/wiki/Generalized_additive_model) doesn't force the predictors to a square as in polynomial regression, but GAMes tries to do a smooth line.
The data we use here is biocapacity of different countries.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
library(psych)
eco <- read.csv("C:/07 - R Website/dataset/ML/biocap.csv")

pairs.panels(eco, 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = FALSE # show correlation ellipses
             )

```

From the scatterplot above, we can see we have quite curve relationship between [**Human Development Index HDI**](https://en.wikipedia.org/wiki/Human_Development_Index) and [**Gross domestic product GDP**](https://en.wikipedia.org/wiki/Gross_domestic_product). Now we can try to build a Generalized Addictive Model with **BiocapacityT** as response variable.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
library(mgcv)
# GAM model
mod_lm = gam(BiocapacityT ~ Population+HDI+Grazing.Footprint+Carbon.Footprint+
                Cropland+Forest.Land+Urban.Land+GDP, data=eco)

                summary(mod_lm)
                # concurvity(mod_lm)

```

From the summary results above, we can see tht GDP is not statistically significant. </br>
one aspect with GAMs is the [**Concurvity**](https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/concurvity.html). It refers to the generalization of collinearity to the GAM setting. </br>
In this case it refers to the situation where a smooth term can be approximated by some combination of the others. </br>
Can lead to unstable estimates. If fact, **Concurvity** can be viewed as a generalization of co-linearity, and causes similar problems of interpretation. We have to drop one of the variable with concurvity.


```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
# Instead of splines specify tensor product smooth
mod_gam3 <- gam(BiocapacityT ~ te(Grazing.Footprint, Cropland, Forest.Land), data=eco)
concurvity(mod_gam3)

vis.gam(mod_gam3, type='response', plot.type='persp',
             phi=30, theta=30,n.grid=500, border=NA)

```
The model **mod_gam3** uses the [**tensor**](https://en.wikipedia.org/wiki/Tensor) smooth for predictors. If we look at the concurvity, in the worst case scenario **Grazing.Footprint** and **Forest.Land** have a strog relationship, but looking at the observed estimation the correlation is not very strong.
The 3D visualization there is aportion that has come down (in red), and we can conclude for some extend that the predictors not contribute to the biocapacoty, and so we have to look to our data into more details.




















